A molecular dynamics study of the solvation of \ce{CO_2}, water, and some organic compounds in the ionic liquids \ce{[emim][B(CN)_4]} and \ce{[emim][NTf_2]}

### 2018-08-02 ionic_liquids.tex

%\documentclass[3p,preprint]{elsarticle}
\documentclass[3p,twocolumn]{elsarticle}
\usepackage[colorlinks,citecolor=blue]{hyperref}
\usepackage{lineno}
\usepackage{amsmath}
\usepackage{ifthen}
\usepackage{commath}
\usepackage{amssymb}
\usepackage{scalerel}
\usepackage{natbib}

@@

\begin{abstract}
In this paper, we calculate solvation free enegies in ionic liquids and predict macroscopic properties required in the design of chemical processes. Carbon dioxide produced in combustion processes is one of the main sources of anthropogenic generation of greenhouse gases. On the other hand, the potential of \ce{[emim][B(CN)_4]} for its capture has recently been reported. For comparison purposes, the solvents \ce{[emim][B(CN)_4]} and \ce{[emim][NTf_2]} were analyzed, both in the condition of infinite dilution and at high concentrations. The simulations confirmed what was observed experimentally, namely the pressure required to obtain the same molality of \ce{CO_2} is notably lower in \ce{[emim][B(CN)_4]}. The infinite-dilution activity coefficients of hexane, benzene, cyclohexane, ethanol, and water in \ce{[emim][B(CN)_4]} were also calculated, and a comparison was made with experimental values recently reported in literature.
\end{abstract}

\begin{keyword}
ionic liquids \sep solvation free energy \sep  \ce{CO_2} 
\end{keyword}

@@

The conventional process for \ce{CO_2} removal comprises a chemical absorption followed by stripping at high temperatures. The most commonly used solvents are volatile organic compounds, specially aqueous solutions of monoethanolamine (MEA). Despite its high absorption capability, the use of MEA suffers from a serious drawback due to the high cost of solvent recovery \cite{Merkel_2010}. Other disadvantages include corrosion and degradation by oxidation, solvent loss by evaporation, which also contributes to increase the environment pollution and generates extra costs for replacing the solvent.

@@

Recently, the potentiality of ILs for \ce{CO_2} capture has been called into question by Carvalho \textit{et al.} \cite{Carvalho_2016}, who warn that the use of molar fraction to express solubility has, to some extent, perpetuated misconceptions regarding the supposed absorption capability of certain ILs. Although most thermodynamic models are formulated in terms of molar fraction, a key factor that dictates the absorption process efficiency is given by the amount of \ce{CO_2} that can be dissolved in a certain \textit{volume} or \textit{mass} of solvent. By expressing the solubility values in terms of molality and performing an exhaustive comparison between various ILs, Ref. \cite{Carvalho_2016} shows that most systems which promote a physical absorption of \ce{CO_2} exhibit a solubility similar or even lower than that observed for the heavy alkane eicosane. This is true for the ``popular'' \ce{[bmim][NTf_2]}, \ce{[bmim][PF_6]}, \ce{[bmim][BF_4]}, as well as for those synthesized from the fluorination of the cation or the anion, such as \ce{[(C_2H_2F_2)mim][NTf_2]} and \ce{[hmim][pFAP]}; the only exception being the family of ILs containing the anion tetracyanoborate \ce{[B(CN)_4]^-}.

There is no obvious explanation accounting for the enhanced solubility promoted by the presence of the anion \ce{[B(CN)_4]^-}. In this sense, molecular simulation provides with a powerful tool for studying solvation processes. To the best of our knowledge, only one study in literature reports the free energy of solvation of \ce{CO_2} in \ce{[emim][B(CN)_4]} \cite{Liu_2014_1} which, in addition, is overestimated. This motivated us to evaluate different force fields of \ce{[emim][B(CN)_4]} and, as we will show below, it was possible to select a model that successfully reproduces the experimental value of the Henry constant of \ce{CO_2}. For comparison, \ce{[emim][B(CN)_4]} and \ce{[emim][NTf_2]} were analyzed, both in the condition of infinite dilution and at high concentrations. The simulations confirmed what is observed in experiments: for \ce{[emim][B(CN)_4]}, a considerably lower pressure is required to dissolve the same amount of \ce{CO_2}. The infinite-dilution activity coefficients of solutes hexane, benzene, cyclohexane, ethanol, and water in \ce{[emim][B(CN)_4]} were also calculated, and a comparison was made with experimental values recently reported in literature.

@@

In the following, the subscripts ``$s$'' and ``$\text{IL}$'' refer to the solute and the solvent (ionic liquid), respectively. For a system containing $N_{\text{IL}}$ solvent molecules and $N_{\text{s}}$ solute molecules at fixed pressure (P) and temperature (T), the difference in free energy between the states associated to $\lambda$ = 1 and $\lambda$ = 0 is formally defined as
\begin{equation}
\Delta G_{\text{sim}} \equiv k_b T \ln \frac{\bigtriangleup (N_s^l,N_{\text{IL}},P,T,\lambda = 1)}{\bigtriangleup (N_s^l,N_{\text{IL}},P,T,\lambda = 0)}, 
\end{equation}
where superscript ``$l$'' refers to the liquid solution, $\bigtriangleup$ denotes the isobaric-isothermal partition function, and $k_b$ is the Boltzmann's constant. Shirts \textit{et al.} \cite{Shirts_2003} rigorously derived the working expression that relates $\Delta G_{\text{sim}}$ to what is experimentally known as the solvation free energy. They found out that
\begin{equation}
\begin{split}
\label{eq:free_solv}
 \Delta G_{\text{solv}}& =  k_b T \ln \left( \frac{P_s}{c_s^{\ast} k_b T} \right)\\ &= \Delta G_{\text{sim}} - k_bT \ln \left( \frac{V^{\ast}}{V_1} \right) ,
\end{split}
\end{equation}
where $V^{\ast}$ and $V_1$ refer to the mean volumes of the systems specified by ($N_s^l-1$,$N_{\text{LI}}$,P,T) and ($N_s^l$,$N_{\text{LI}}$,P,T), respectively; $c_s^{\ast}$ = $N_s^l/V_1$ is the numerical concentration of the solute in the liquid phase and $P_s$ is the partial pressure of the solute in the gas phase. It should be noted that in the derivation of Eq.~\eqref{eq:free_solv}, the gas phase was treated as an ideal gas.

@@

For the ILs of interest to the present investigation, high concentrations of \ce{CO_2} are found at pressure values for which, in principle, the ideal gas assumption may be not valid. In this case, we shall correct the ideal gas contribution to the chemical potential, as follows
\begin{equation}
\label{eq:mu_gas_real}
\mu^{g}_s = \mu^{ig}_s + RT \ln \phi_s,
\end{equation}
where $\phi_s$ is the fugacity coefficient of the solute. This equation, together with the fact that $f^{g}_s$ = $P_s \phi_s$, are used to evaluate $\mu^{l}_s$ = $\mu^{g}_s$, which yields
\begin{equation}
\label{eq:fgas_d}
f^{g}_s = \exp  \left( \beta \Delta G_{\text{sim}} \right) +  k_b T c_s^{\ast}, 
\end{equation}
where $f^{g}_s$ is the fugacity of the solute in the gas phase. Having calculated $f^{g}_s$, $P_s$ can be predicted using the following equation
\begin{equation}
\label{eq:fgas}
f^{g}_s = P_s \exp\left[ \frac{B_s(T) P_s}{R T} \right],
\end{equation}
where $B_s$ corresponds to the second virial coefficient. Here we employ the correlation of $B_s$ for \ce{CO_2} introduced by Holste \textit{et al.} \cite{Holste_1987}, which is
\begin{equation}
B_s = B_0 + \frac{B_1}{T} + \frac{B_2}{T^2},
\end{equation}
where $B_0$ = 23.02991 cm$^3$/mol, $B_1$ = $-2.455297$ $\times$ 10$^3$ (Kcm$^3$/mol) and $B_2$ = -1.22675$\times$10$^7$ (K$^2$cm$^3$/mol).

As supported by preliminary simulations, it is reasonable to assume $\Delta G_{\text{sim}}$ as independent of pressure over the range of experimental data considered in this work. Then, the equilibrium pressure is determined by solving Eq.~\eqref{eq:fgas} for $P_s$.

@@

The solutes considered in this work are hexane, benzene, \ce{CO_2}, ethanol, cyclohexane and water. Hexane and benzene were modeled using the OPLS-AA force field \cite{Jorgensen_1996}, while the TRAPPE potential \cite{Chen_2001,Potoff_2001} was employed for ethanol and \ce{CO_2}. Cyclohexane was simulated using a rigid-body model \cite{munoz2015lennard} whose sites only interact via a LJ potential. For water, we employ the TIP4P/2005 model \cite{Abascal_2005,Vega_2011}.

@@

With the aim of setting up a computationally efficient protocol for the ILs, in this work we assess the possibility of modeling the ions as rigid bodies and employing the damped shifted force (DSF) method \cite{Fennell2006} for computing the electrostatic interactions. As an alternative to the Ewald schemes \cite{Ewald_1921,Darden_1993,Hockney_1988}, McCann \textit{et al.} \cite{McCann_2013} tested the DSF method for various ILs modeled with the OPLS-AA force fields. They concluded that using $\alpha$ = 0.2 {\AA}$^{-1}$ and a cutoff distance ($r_c$) of 15 {\AA} provide the best energetic agreement with the Ewald method. In this work, the comparisons are made in terms of the simulated densities and radial distribution functions for which, as will be shown in the next section, $\alpha$ = 0.2 {\AA}$^{-1}$ and $r_c$ = 12 {\AA} resulted in a good agreement. These values were also employed for the charged solutes, with the exception of hexane, for which we applied undamped Coulombic interactions improving the prediction of the chemical potential.

The NPT algorithm of Kamberaj \textit{et al.} \cite{Kamberaj_2005} was used to perform the simulations with rigid bodies. For flexible molecules, we applied the scheme due to Shinoda \textit{et al.} \cite{Shinoda2004} C-H bonds were constrained by the SHAKE method \cite{Ryckaert1977}. For both integrators, we used a time-step size ($\Delta t$) of 1 fs.

@@

In this work, we carried out independent simulation at each state of $\lambda_{\text{Coul}}$, while the expanded ensemble (EE) method~\cite{Lyubartsev_1992} was applied for sampling the $\lambda_{\,\text{vdW}}$ set. In short, an EE is constructed as a weighted sum of subensembles, each assuming a different value of the variable chosen to explore the free energy landscape, which in our case is $\lambda_{\,\text{vdW}}$. This way, the sampling of an EE comprises both sampling each subensamble with fixed variables (N,P,T,$\lambda_{\,\text{vdW}}$) and transitions in $\lambda_{\,\text{vdW}}$. Thus, a remarkable feature of the EE method is the possibility of exploring the whole alchemical route in a single simulation. In this work, we perform a sequential sampling of the EE \cite{Chodera_2011_2}, which consists in sampling each subensamble via NPT molecular dynamics (MD) for a determined number of steps $N^{\text{EE}}$ followed by an attemp to change $\lambda_{\,\text{vdW}}$ using MC, and so on. This change in $\lambda_{\,\text{vdW}}$ is randomly proposed according to an independent sampling scheme \cite{Chodera_2011_2}. In addition, the set of (arbitrary) weights $\eta$ of the subensambles enters in the acceptance criteria of MC, so it is possible to tune $\eta$ in order to enhance the sampling. If differences $\eta_i$ - $\eta_j$ were closed to $\Delta G_{ji}$, we expect an uniform sampling in $\lambda_{\,\text{vdW}}$ \cite{Lyubartsev_1992}. Of course, this can be achieved only iteratively. The EE procedure, together with the \textit{soft-core} potential and the DSF method with linear coupling were implemented in the LAMMPS software package \cite{Plimpton1995}, which we used for carrying out all the simulations.

@@

The initial configurations were obtained with PLAYMOL \cite{playmol}, a software integrated with PACKMOL \cite{Mart_nez_2009}, using the tool \textit{build} to specify geometric data (bond and angle lengths, and dihedral angles), from which PLAYMOL automatically determines the Cartesian coordinates. We first carried out NVT MD simulations of the isolated ion pairs in order to obtain a minimum energy conformation. For this, we used the Nos\'{e}-Hoover thermostat \cite{Martyna1992} and a cubic box of length L = 50 {\AA} with non-periodic boundary conditions. After a short run with $\Delta t$ = 0.5 fs, we carried out a ``quenching'' simulation with $\Delta t$ = 1 fs, in which the temperature is gradually reduced until reaching a value close to 0 K. The resulting coordinates were then specified in PLAYMOL \cite{playmol} to build the configurations of the liquid phases.

@@

In this section, we present results for the following two cases: (1) flexible ions and the Particle-Particle-Particle-Mesh (PPPM) method~\cite{Hockney_1988} with a tolerance of 1$\times$10$^{-4}$, and (2) ions modeled as rigid bodies with the DSF method \cite{Fennell2006} ($\alpha$ = 0.2 {\AA}$^{-1}$) for calculating the Coulombic interactions.
In both cases, $r_c$ = 12 {\AA}.

The configurations of rigid bodies were obtained after carrying out an analysis of key dihedrals, whose flexibility could lead to significant conformational changes in \ce{[emim]^+} and \ce{[NTf_2]^-}.
On the other hand, small and highly symmetric anions, such as \ce{PF_6} and \ce{BF_4} are usually modeled as rigid bodies \cite{HANKE_2001,Hanke_2003,Lynden_Bell_2006}, which justifies the strategy in the case of \ce{[B(CN)_4]^-}.
The nomenclature used in this paper is shown in Fig~\ref{fig:atoms_id}.
In the \textit{united-atoms} version of \ce{[emim]^+}, the hydrogens ranging from H4 to H11 are eliminated.

\begin{figure}[H]
\centering
\includegraphics[width=\linewidth]{ions_paper_test.pdf}
\caption{Adopted nomenclature for 1-ethyl-3-methylimidazolium  (\ce{[emim]^+}), tetracyano borate (\ce{[B(CN)_4]^-}) and bis(trifluoromethylsulfonyl)amide (\ce{[NTf_2]^-}).}
\label{fig:atoms_id}
\end{figure}

Using the software TRAVIS \cite{Brehm_2011}, we obtained the combined distribution functions (CDFs) of the dihedral angles of C1-N2-C5-C6 and C2-N2-C5-C6, as sampled in NPT-MD simulations of 250 ion pairs at $T = 298.0~\text{K}$ and $P = 1.0~\text{bar}$.
In Fig.~\ref{fig:dihedrals_emim} we show the CDFs for each force field of \ce{[emim][B(CN)_4]}.
A first inspection of this figure reveals a similar behavior of the models due to Koller \textit{et al.} \cite{Koller_2012} and Liu \textit{et al.} \cite{Liu_2014}.
This was expected as both are based on the AMBER framework \cite{Cornell_1995}.
In addition, Figs.~\ref{fig:dihedrals_emim}(a) and \ref{fig:dihedrals_emim}(b) are symmetric, which means that atoms C1, N2, and C2 are coplanar.
In Fig.~\ref{fig:dihedrals_emim}(a), the most likely angles can be found in a narrow region around 90$^{\circ}$ and 270$^{\circ}$.
Since these angles correspond to an equivalent configuration in the \textit{united-atom} \ce{[emim]^+}, we built all the rigid bodies specifying 90$^{\circ}$ for the dihedral angle of C6-C5-N2-C1.
The other models show a larger flexibility with high probability regions that extend in a broad range of angle values.
Despite this, it was possible to build more than one type of rigid body by specifying the most probables angles of C6-C5-N2-C1.
In the case of Liu \textit{et al.} \cite{Liu_2014}, these angles correspond to 90$^{\circ}$ and -90$^{\circ}$, while for the model of Batista \textit{et al.} \cite{Batista_2015}, we employed 102.6$^{\circ}$ and -102.6$^{\circ}$.
In these cases, two types of rigid bodies were built in equal proportion.
Fig.~\ref{fig:dihedrals_emim}(d) shows that the model of Weber and Kirchner \cite{Weber_2016} there is a non-negligible probability of sampling the angle of 0$^{\circ}$.
In this case, we defined four types of rigid bodies: 150$^{\circ}$, 180$^{\circ}$ and 210$^{\circ}$ were employed to build 30~\%, 25~\% and 30~\% of the rigid bodies, respectively; and for the remaining 15~\% we specified the angle of 0$^{\circ}$.

@@

A study of dihedrals was also carried out for \ce{[emim][NTf_2]}.
The CDF of dihedrals angles of C1-N2-C5-C6 and C2-N2-C5-C6, shown in Fig.~\ref{fig:die_ntf2}(a), resembles the results of Fig.~\ref{fig:dihedrals_emim}(d), but the most likely angle of C1-N2-C5-C6 is 0$^{\circ}$.
This similarity was expected as both models of \ce{[emim]^+} are based on the force field developed by Canongia-Lopes and P\'{a}dua~ \cite{Canongia_Lopes_2006}.
With regards to the anion \ce{[NTf_2]^-}, Deetlefs \textit{et al.} \cite{Deetlefs_2006} have experimentally determined the conformer population of the liquid phase and observed that \ce{[NTf_2]^-} is mostly in its \textit{trans} conformation.
Fig.~\ref{fig:die_ntf2}(b), which corresponds to the CDF for the contiguous C1-S1-N1-S2 and S1-N1-S2-C2 chains, qualitatively verifies the observation of Deetlefs \textit{et al.} \cite{Deetlefs_2006}.
Given the considerable flexibility of \ce{[NTf_2]^-}, a more rigorous strategy is required for building the rigid bodies.
Thus, the simulations of \ce{[emim][NTf_2]} were performed with flexible ions and the DSF method, which results in a computed density of 1.495 $\pm$ 0.009 g/cm$^3$ with an error of 1.12\% with respect to the experimental value~\cite{Tokuda_2005}.

\begin{figure}[]
\centering
  \includegraphics[width=\linewidth]{Ludwig.png}%

@@

\subsection{Henry constant of \ce{CO_2} in \ce{[emim][B(CN)_4]} and [emim][Tf\textsubscript{2}N]}
\label{sec:henry_results}

>>>>

%\documentclass[3p,preprint]{elsarticle}
\documentclass[3p,twocolumn]{elsarticle}
\usepackage[colorlinks,citecolor=blue]{hyperref}
\usepackage{lineno}
\usepackage{amsmath}
\usepackage{ifthen}
\usepackage{commath}
\usepackage{amssymb}
\usepackage{scalerel}
\usepackage{natbib}
\usepackage[version=4]{mhchem}

@@

\begin{abstract}
In this paper, we calculate solvation free energies in ionic liquids and predict macroscopic properties required in the design of chemical processes.
Carbon dioxide from combustion processes is one of the main sources of anthropogenic generation of greenhouse gases.
On the other hand, the potential of \ce{[emim][B(CN)_4]} for its capture has recently been reported.
For comparison, the solvents \ce{[emim][B(CN)_4]} and \ce{[emim][NTf_2]} are analyzed, both at infinite dilution and at high concentrations of \ce{CO_2}.
Our results confirm what is experimentally observed, namely a considerably lower pressure is required to dissolve a certain amount of \ce{CO_2} in \ce{[emim][B(CN)_4]}.
The infinite-dilution activity coefficients of hexane, benzene, cyclohexane, ethanol, and water in \ce{[emim][B(CN)_4]} are also calculated, and a comparison is made with experimental values recently reported in literature.
\end{abstract}

\begin{keyword}
ionic liquids \sep solvation free energy \sep  \ce{CO_2} Henry constant \sep infinite-dilution activity coefficient 
\end{keyword}

@@

The conventional process for \ce{CO_2} removal comprises a chemical absorption followed by stripping at high temperatures.
The most commonly used solvents are volatile organic compounds, specially aqueous solutions of monoethanolamine (MEA).
Despite its high absorption capability, the use of MEA suffers from a serious drawback due to the high cost of solvent recovery \cite{Merkel_2010}.
Other disadvantages include corrosion and degradation by oxidation, solvent loss by evaporation, which also contributes to increase the environmental pollution and generates extra costs for replacing the solvent.

@@

Recently, the potentiality of ILs for \ce{CO_2} capture has been called into question by Carvalho \textit{et al.} \cite{Carvalho_2016}, who warn that the use of molar fraction to express solubility has, to some extent, perpetuated misconceptions regarding the supposed absorption capability of certain ILs.
Although most thermodynamic models are formulated in terms of molar fraction, a key factor that dictates the absorption process efficiency is given by the amount of \ce{CO_2} that can be dissolved in a certain \textit{volume} or \textit{mass} of solvent.
By expressing the solubility values in terms of molality and performing an exhaustive comparison between various ILs, Ref. \cite{Carvalho_2016} shows that most systems which promote a physical absorption of \ce{CO_2} exhibit a solubility similar, or even lower, than that observed for the heavy alkane eicosane.
This is true for the ``popular'' \ce{[bmim][NTf_2]}, \ce{[bmim][PF_6]}, \ce{[bmim][BF_4]}, as well as for those synthesized from the fluorination of the cation or the anion, such as \ce{[(C_2H_2F_2)mim][NTf_2]} and \ce{[hmim][pFAP]}; the only exception being the family of ILs containing the anion tetracyanoborate \ce{[B(CN)_4]^-}.

There is no obvious explanation accounting for the enhanced solubility promoted by the presence of \ce{[B(CN)_4]^-}.
In this context, molecular simulation provides with a powerful tool for studying solvation processes.
To the best of our knowledge, only one study in literature reports the free energy of solvation of \ce{CO_2} in \ce{[emim][B(CN)_4]} \cite{Liu_2014_1} which, in addition, is overestimated.
This motivated us to evaluate different force fields of \ce{[emim][B(CN)_4]} and, as we will show below, it was possible to find a model that successfully reproduces the experimental value of the Henry constant of \ce{CO_2}.
For comparison, \ce{[emim][B(CN)_4]} and \ce{[emim][NTf_2]} were analyzed, both at infinite dilution and at high concentrations of \ce{CO_2}.
The simulations confirmed what is observed in experiments: for \ce{[emim][B(CN)_4]}, a considerably lower pressure is required to dissolve the same amount of \ce{CO_2}.
The infinite-dilution activity coefficients of solutes hexane, benzene, cyclohexane, ethanol, and water in \ce{[emim][B(CN)_4]} were also calculated, and a comparison was made with experimental values recently reported in literature.

@@

In the following, the subscripts ``$s$'' and ``$\text{IL}$'' refer to the solute and the solvent (ionic liquid), respectively.
For a system containing $N_{\text{IL}}$ solvent molecules and $N_{\text{s}}$ solute molecules at fixed pressure (P) and temperature (T), the difference in free energy between the states associated to $\lambda$ = 1 and $\lambda$ = 0 is formally defined as
\begin{equation}
\Delta G_{\text{sim}} \equiv k_b T \ln \frac{\bigtriangleup (N_s^l,N_{\text{IL}},P,T,\lambda = 1)}{\bigtriangleup (N_s^l,N_{\text{IL}},P,T,\lambda = 0)}, 
\end{equation}
where superscript ``$l$'' refers to the liquid solution, $\bigtriangleup$ denotes the isobaric-isothermal partition function, and $k_b$ is the Boltzmann's constant.
Shirts \textit{et al.} \cite{Shirts_2003} rigorously derived the working expression that relates $\Delta G_{\text{sim}}$ to what is experimentally known as the solvation free energy.
They found out that
\begin{equation}
\begin{split}
\label{eq:free_solv}
 \Delta G_{\text{solv}}& =  k_b T \ln \left( \frac{P_s}{c_s^{\ast} k_b T} \right)\\ &= \Delta G_{\text{sim}} - k_bT \ln \left( \frac{V^{\ast}}{V_1} \right),
\end{split}
\end{equation}
where $V^{\ast}$ and $V_1$ refer to the mean volumes of the systems specified by ($N_s^l-1$,$N_{\text{LI}}$,P,T) and ($N_s^l$,$N_{\text{LI}}$,P,T), respectively; $c_s^{\ast}$ = $N_s^l/V_1$ is the numerical concentration of the solute in the liquid phase and $P_s$ is the partial pressure of the solute in the gas phase.
It should be noted that in the derivation of Eq.~\eqref{eq:free_solv}, the gas phase is treated as an ideal gas.

@@

For the ILs of interest to the present investigation, high concentrations of \ce{CO_2} are found at pressure values for which, in principle, the ideal gas assumption may be not valid.
In this case, we shall correct the ideal gas contribution to the chemical potential, as follows
\begin{equation}
\label{eq:mu_gas_real}
\mu^{g}_s = \mu^{ig}_s + RT \ln \phi_s,
\end{equation}
where $\phi_s$ is the fugacity coefficient of the solute.
This equation, together with the fact that $f^{g}_s$ = $P_s \phi_s$, are used to evaluate $\mu^{l}_s$ = $\mu^{g}_s$, which yields
\begin{equation}
\label{eq:fgas_d}
f^{g}_s = \exp  \left( \beta \Delta G_{\text{sim}} \right) +  k_b T c_s^{\ast}, 
\end{equation}
where $f^{g}_s$ is the fugacity of the solute in the gas phase.
Having calculated $f^{g}_s$, $P_s$ can be predicted using the following equation
\begin{equation}
\label{eq:fgas}
f^{g}_s = P_s \exp\left[ \frac{B_s(T) P_s}{R T} \right],
\end{equation}
where $B_s$ corresponds to the second virial coefficient.
Here, we employ the correlation of $B_s$ for \ce{CO_2} introduced by Holste \textit{et al.} \cite{Holste_1987}, which is
\begin{equation}
B_s = B_0 + \frac{B_1}{T} + \frac{B_2}{T^2},
\end{equation}
where $B_0$ = 23.02991 cm$^3$/mol, $B_1$ = -2.455297 $\times$ 10$^3$ (Kcm$^3$/mol) and $B_2$ = -1.22675$\times$10$^7$ (K$^2$cm$^3$/mol).

As supported by preliminary simulations, it seems reasonable to assume $\Delta G_{\text{sim}}$ as independent of pressure over the range of experimental data considered in this work.
Then, the equilibrium pressure is determined by directly solving Eq.~\eqref{eq:fgas} for $P_s$.

@@

The solutes considered in this work are hexane, benzene, \ce{CO_2}, ethanol, cyclohexane and water.
Hexane and benzene were modeled using the OPLS-AA force field \cite{Jorgensen_1996}, while the TRAPPE potential \cite{Chen_2001,Potoff_2001} was employed for ethanol and \ce{CO_2}.
Cyclohexane was simulated using a rigid-body model \cite{munoz2015lennard} whose sites only interact via a LJ potential
For water, we employ the TIP4P/2005 model \cite{Abascal_2005,Vega_2011}.

@@

With the aim of setting up a computationally efficient protocol for the ILs, in this work we assess the possibility of modeling the ions as rigid bodies and employing the damped shifted force (DSF) method \cite{Fennell2006} for computing the electrostatic interactions.
As an alternative to the Ewald schemes \cite{Ewald_1921,Darden_1993,Hockney_1988}, McCann \textit{et al.} \cite{McCann_2013} tested the DSF method for various ILs modeled with the OPLS-AA force fields.
They concluded that using $\alpha$ = 0.2 {\AA}$^{-1}$ and a cutoff distance ($r_c$) of 15 {\AA} provide the best energetic agreement with the Ewald method.
In this work, comparisons are made in terms of the simulated densities and radial distribution functions for which, as will be shown in the next section, $\alpha$ = 0.2 {\AA}$^{-1}$ and $r_c$ = 12 {\AA} resulted in a good agreement.
These values were also employed for the charged solutes, with the exception of hexane, for which we applied undamped Coulombic interactions improving the prediction of the chemical potential.

The NPT algorithm of Kamberaj \textit{et al.} \cite{Kamberaj_2005} was used to perform the simulations with rigid bodies.
For flexible molecules, we applied the scheme due to Shinoda \textit{et al.} \cite{Shinoda2004} and C-H bonds were constrained by the SHAKE method \cite{Ryckaert1977}.
For both integrators, we used a time-step size ($\Delta t$) of 1 fs.

@@

In this work, we carried out independent simulation at each state of $\lambda_{\text{Coul}}$, while the expanded ensemble (EE) method~\cite{Lyubartsev_1992} was applied for sampling the $\lambda_{\,\text{vdW}}$ set.
In short, an EE is constructed as a weighted sum of subensembles, each assuming a different value of the variable chosen to explore the free energy landscape, which in our case is $\lambda_{\,\text{vdW}}$.
This way, the sampling of an EE comprises both sampling each subensamble with fixed variables (N,P,T,$\lambda_{\,\text{vdW}}$) and transitions in $\lambda_{\,\text{vdW}}$.
Thus, a remarkable feature of the EE method is the possibility of exploring the whole alchemical route in a single simulation.
In this work, we perform a sequential sampling of the EE \cite{Chodera_2011_2}, which consists in sampling each subensamble via NPT molecular dynamics (MD) for a determined number of steps $N^{\text{EE}}$ followed by an attempt to change $\lambda_{\,\text{vdW}}$ using MC, and so on.
The $\lambda_{\,\text{vdW}}$ transitions are performed according to an independent sampling scheme \cite{Chodera_2011_2}.
In addition, the set of (arbitrary) weights $\eta$ of the subensambles enters in the acceptance criteria of MC, so it is possible to tune $\eta$ in order to enhance the sampling.
If differences $\eta_i$ - $\eta_j$ were closed to $\Delta G_{ji}$, we expect an uniform sampling in $\lambda_{\,\text{vdW}}$ \cite{Lyubartsev_1992}.
Of course, this can be achieved only iteratively.
The EE procedure, together with the \textit{soft-core} potential and the DSF method with linear coupling were implemented in the LAMMPS software package \cite{Plimpton1995}, which we used for carrying out all the simulations.

@@

The initial configurations were obtained with PLAYMOL \cite{playmol}, a software integrated with PACKMOL \cite{Mart_nez_2009}, using the tool \textit{build} to specify geometric data (bond and angle lengths, and dihedral angles), from which PLAYMOL automatically determines the Cartesian coordinates.
We first carried out NVT MD simulations of the isolated ion pairs in order to obtain a minimum energy conformation.
For this, we used the Nos\'{e}-Hoover thermostat \cite{Martyna1992} and a cubic box of length L = 50 {\AA} with non-periodic boundary conditions.
After a short run with $\Delta t$ = $0.5~/text{fs}$, we carried out a ``quenching'' simulation with $\Delta t$ = $1 ~/text{fs}$, in which the temperature is gradually reduced until reaching a value close to $0~/text{K}$.
The resulting coordinates were then specified in PLAYMOL \cite{playmol} to build the configurations of the liquid phases.

@@

In this section, we present results for the following two cases: (1) flexible ions together with the Particle-Particle-Particle-Mesh (PPPM) method~\cite{Hockney_1988} with a tolerance of 1$\times$10$^{-4}$ for calculating the Coulombic interactions, and (2) ions modeled as rigid bodies with the DSF method \cite{Fennell2006} ($\alpha$ = 0.2 {\AA}$^{-1}$) as the alternative to the PPPM scheme.
In both cases, $r_c$ = 12 {\AA}.
The nomenclature used in this paper is shown in Fig~\ref{fig:atoms_id}.
In the \textit{united-atoms} version of \ce{[emim]^+}, hydrogens ranging from H4 to H11 are eliminated.

The rigid bodies were built after carrying out an analysis of key dihedrals, whose flexibility could lead to significant conformational changes in \ce{[emim]^+} and \ce{[NTf_2]^-}.
On the other hand, small and highly symmetric anions, such as \ce{PF_6} and \ce{BF_4} are usually modeled as rigid bodies \cite{HANKE_2001,Hanke_2003,Lynden_Bell_2006}, which justifies the strategy in the case of \ce{[B(CN)_4]^-}.

\begin{figure}[ht]
\centering
\includegraphics[width=\linewidth]{ions_paper.pdf}
\caption{Adopted nomenclature for 1-ethyl-3-methylimidazolium  (\ce{[emim]^+}), tetracyano borate (\ce{[B(CN)_4]^-}) and bis(trifluoromethylsulfonyl)amide (\ce{[NTf_2]^-}).}
\label{fig:atoms_id}
\end{figure}

Using the software TRAVIS \cite{Brehm_2011}, we obtained the combined distribution functions (CDFs) of the dihedral angles of C1-N2-C5-C6 and C2-N2-C5-C6, as sampled in NPT-MD simulations of 250 ion pairs at $T = 298.0~\text{K}$ and $P = 1.0~\text{bar}$.
In Fig.~\ref{fig:dihedrals_emim} we show the CDFs for each force field of \ce{[emim][B(CN)_4]}.
A first inspection of this figure reveals a similar behavior of the models due to Koller \textit{et al.} \cite{Koller_2012} and Liu \textit{et al.} \cite{Liu_2014}.
This was expected as both are based on the AMBER framework \cite{Cornell_1995}.
In addition, Figs.~\ref{fig:dihedrals_emim}(a) and \ref{fig:dihedrals_emim}(b) are symmetric, which means that atoms C1, N2, and C2 are coplanar.
In Fig.~\ref{fig:dihedrals_emim}(a), the most likely angles can be found in a narrow region around 90$^{\circ}$ and 270$^{\circ}$.
Since these angles correspond to an equivalent configuration in the \textit{united-atom} \ce{[emim]^+}, in the case of Koller \textit{et al.} \cite{Koller_2012} we built all the rigid bodies by specifying 90$^{\circ}$ for the dihedral angle of C6-C5-N2-C1.
The other models show a larger flexibility with high probability regions that extend in a broad range of angle values.
Despite this, it was possible to build more than one type of rigid body by specifying the most probables angles of C6-C5-N2-C1.
In the case of Liu \textit{et al.} \cite{Liu_2014}, these angles correspond to 90$^{\circ}$ and -90$^{\circ}$, while for the model of Batista \textit{et al.} \cite{Batista_2015}, we employed 102.6$^{\circ}$ and -102.6$^{\circ}$.
These two types of rigid bodies were built in equal proportion.
In Fig.~\ref{fig:dihedrals_emim}(d), which corresponds to Weber and Kirchner \cite{Weber_2016}, one sees that the angle As this is the most flexible model, we defined four types of rigid bodies: 150$^{\circ}$, 180$^{\circ}$ and 210$^{\circ}$ were employed to build 30~\%, 25~\% and 30~\% of them, respectively; and 0$^{\circ}$ for the remaining 15~\%.

@@

A study of dihedrals was also carried out for \ce{[emim][NTf_2]}.
The CDF of the dihedrals angles of C1-N2-C5-C6 and C2-N2-C5-C6, shown in Fig.~\ref{fig:die_ntf2}(a), resembles the results of Fig.~\ref{fig:dihedrals_emim}(d), but the most likely angle of C1-N2-C5-C6 is 0$^{\circ}$.
This similarity was expected as both models of \ce{[emim]^+} are based on the force field developed by Canongia-Lopes and P\'{a}dua~ \cite{Canongia_Lopes_2006}.
With regards to the anion \ce{[NTf_2]^-}, Deetlefs \textit{et al.} \cite{Deetlefs_2006} have experimentally determined the conformer population of the liquid phase and observed that \ce{[NTf_2]^-} is mostly in its \textit{trans} conformation.
Fig.~\ref{fig:die_ntf2}(b), which corresponds to the CDF for the contiguous C1-S1-N1-S2 and S1-N1-S2-C2 chains, qualitatively verifies the observation of Deetlefs \textit{et al.} \cite{Deetlefs_2006}.
Given the considerable flexibility of \ce{[NTf_2]^-}, a more rigorous strategy is required for building the rigid bodies.
Thus, the simulations of \ce{[emim][NTf_2]} were performed considering flexible ions and the DSF method, which results in a computed density of 1.495 $\pm$ 0.009 g/cm$^3$ with an error of 1.12\% with respect to the experimental value~\cite{Tokuda_2005}.

\begin{figure}[ht]
\centering
  \includegraphics[width=\linewidth]{Ludwig.png}%

@@

\subsection{Henry constant of \ce{CO_2} in \ce{[emim][B(CN)_4]} and \ce{[emim][NTf_2]}}
\label{sec:henry_results}

In Table \ref{table:henry} we present the results of solvation free energies of \ce{CO_2} along with the Henry constants ($K_s$) obtained by means of Eq.~\eqref{eq:henry_eq}.
Table \ref{table:henry} also includes the results for [emim][Tf\textsubscript{2}N] as well as the experimental $K_s$ reported by Mahurin \textit{et al.} \cite{Mahurin_2010} and Finotello \textit{et al.} \cite{Finotello_2008}.
The uncertainties in $K_s$ were calculated using standard error propagation.
